Results

### 2016-01-28 doc/paper/results.Rmd
a little more results

There are two major components to the accuracy of a variant caller. The first, sensitivity, describes how well a variant caller can identify variants that are know to be real. In other words what proportion of the true positives are found.  The second, specificity, measures how well a variant caller can correctly identify true negatives. For example, the influenza genome is roughly 13KB. There is potential for 3 variant nucleotides at each position leaving a variant caller with about 39,000 potential variants that can be called. The vast majority of these potential variants will not be present in the starting population and are potential false positives. A variant caller that correctly identifies all potential false positives as true negatives has a specificity of 1.

There is an obvious trade off between sensitivity and specificity. It may be the case that a variant caller has near perfect sensitivity, ie. it finds all the true positives in a sample. However, if in the process it has low specificity and identifies many false positives that will confound down stream analysis then the high sensitivity is all for naught. Conversely, a variant caller may have perfect specificity and identify no false positives; however, if in process it has low sensitivity and identifies only a few true positives again it will have limited applications.

The trade off between sensitivity and specificity can be represented as a reciter operator curve or ROC (`r xamp_roc`). A ROC plots the sensitivity of an assay along the y axis and 1-specificity along the x axis. Therefore a perfect assay with a sensitivity and specificity of 1 will lie in at the coordinates (0,1) in the top left hand corner of the plot.

>>>>

There are two major components to the accuracy of a variant caller. The first, sensitivity, describes how well a variant caller can identify variants that are known to be real. In other words what proportion of the true positives are found. A sensitivity of 1 indicates that all true positives are found. The second, specificity, measures how well a variant caller can correctly identify true negatives. For example, the influenza genome is roughly 13KB. There are 3  potential variant nucleotides at each position leaving a variant caller with about 39,000 potential variants that can be called. The vast majority of these potential variants will not be present in the starting population and are potential false positives. A variant caller that correctly identifies all potential false positives as true negatives has a specificity of 1. A perfect variant caller will have a sensitivity and specificity of 1. It will identify all true variants and only the true variants.

There is an obvious trade off between sensitivity and specificity. It may be the case that a variant caller has near perfect sensitivity, and  can find all the true positives in a sample. However, if in the process it has low specificity and identifies many false positives down stream analysis will be confounded. Conversely, a variant caller may have perfect specificity and identify no false positives; however, if in process it has low sensitivity and identifies only a few true positives again it will have limited applications.

The trade off between sensitivity and specificity can be represented as a receiver operator curve or ROC (`r xamp_roc`). An ROC plots the sensitivity of an assay along the y axis and 1-specificity along the x axis. To make a ROC for a variant caller one must test an assay on known data. The outcomes can then be stratified according to some metric that represents our belief that a given variant is real. Often this is a p value or quality score. Cut offs can then be set (`r xamp_roc` A) and because the true outcomes are known one can calculate the sensitivity and specificity. The curve is then made up of points representing the sensitivity and specificity of the variant caller at various cut offs. A perfect ROC in which all the true positives can be separated from all potential false positives is a right angle that follows the upper left perimeter of the plot and passes through the point (0,1)  (`r xamp_roc` B).
